---
title: "Data Methods Group Project"
author: "Genna Campain"
date: "9/10/2022"
output: html_document
---
# Scrape and edit top songs lists
Load packages
```{r, echo = FALSE}
library(spotifyr)
library(magrittr)
library(geniusr)
library(dplyr)
library(tidyverse)
library(tidytext)
library(textdata)
library(stringr)
```
Access token
```{r}
id <- "9bd81604c1724db4b9ad68f64dd7fbfe"
secret <- "791feacce6c143f5bf2ecb23d517e055"
Sys.setenv(SPOTIFY_CLIENT_ID = id)
Sys.setenv(SPOTIFY_CLIENT_SECRET = secret)
access_token <- get_spotify_access_token()
token <- "QPPNNpRdh3O3T-e7LbSGTeYzIefKe1ZL6CeoR26DKwRxdMbln6audf3TNViNVhEG"
Sys.setenv(GENIUS_API_TOKEN = token)
```
Scraping top 100 songs of the year playlists
```{r}
year_id <- data.frame(id = c("37i9dQZF1DXe2bobNYDtW8", "37i9dQZF1DWVRSukIED0e9", "37i9dQZF1DX7Jl5KP2eZaS", "5GhQiRkGuqzpWZSE7OU4Se", "37i9dQZF1DX18jTM2l2fJY"), year = 2018:2022)

top_songs <- data.frame()
for(j in 1:5) {
playlist_id <- year_id[j, 1]
year <- year_id[j, 2]
tracks <- get_playlist_tracks(
    playlist_id = playlist_id,
    fields = c("track.artists", "track.duration_ms", "track.explicit", "track.id", "track.name", "track.popularity", "track.album.name", "track.album.release.date")
) %>%
  mutate(year = year)
top_songs <- rbind(top_songs, tracks)
}
```
Subset to take top 50 songs by year, unpack artist names
```{r, echo = FALSE}
top_songs <- top_songs %>% 
  group_by(year) %>% 
  mutate(rown = row_number()) %>%
  ungroup()
top_songs <- top_songs[which(top_songs$rown <= 50), ]

artists <- top_songs$track.artists
artists2 <- matrix(0, 0, nrow = 250, ncol = 1)
for(j in 1:250){ 
     df <- artists[[j]]
     name <- df$name
     artists2[j] <- name
}
top_songs <- cbind(top_songs, artists2)
songinfo <- select(top_songs, c(track.name, artists2, year)) %>%
  rename(track.artists = artists2)
```
# Scrape song lyrics
Edit song and artist names to match Genius website
```{r}
songinfo$track.name <- gsub("\\s*\\([^\\)]+\\)","",as.character(songinfo$track.name))
songinfo2018 <- songinfo[which(songinfo$year == 2018), ]
songinfo2019 <- songinfo[which(songinfo$year == 2019), ]
songinfo2020 <- songinfo[which(songinfo$year == 2020), ]
songinfo2021 <- songinfo[which(songinfo$year == 2021), ]
songinfo2022 <- songinfo[which(songinfo$year == 2022), ]


# 2018
songinfo2018[19, 1] <- "LOVE."
songinfo2018[15, 2] <- "Lil Baby and Drake"
songinfo2018[6, 2] <- "Juice wrld"
songinfo2018[13, 1] <- "happier"
songinfo2018[17, 2] <- "Offset and Metro Boomin"
songinfo2018[21, 2] <- "Anne marie"
songinfo2018[32, 2] <- "Dynoro and gigi dagostino"
songinfo2018[33, 2] <- "G Eazy and Halsey"
songinfo2018[33, 1] <- "Him and I"
songinfo2018[36, 1] <- "Te bote"
songinfo2018[39, 1] <- "1 2 3"
songinfo2018[39, 2] <- "Sofia Reyes"
songinfo2018[41, 2] <- "Hailee Steinfeld and Alesso"
songinfo2018[48, 1] <- "Dejala Que Vuelva"
# 2019
songinfo2019[3, 2] <- "Shawn Mendes and Camila Cabello"
songinfo2019[3, 1] <- "Senorita"
songinfo2019[5, 1] <- "Sunflower"
songinfo2019[44, 1] <- "10000 hours"
songinfo2019[44, 2] <- "Dan shay"
# 2020
songinfo2020[21, 2] <- "Jawsh 685 and Jason Derulo"
songinfo2020[21, 1] <- "Savage love laxed siren beat"
songinfo2020[26, 1] <- "Senorita"
songinfo2020[26, 2] <- "Shawn Mendes and Camila Cabello"
songinfo2020[38, 1] <- "Sunflower"
songinfo2020[39, 1] <- "Hawai"
songinfo2020[42, 1] <- "ritmo bad boys for life"
songinfo2020[42, 2] <- "The black eyed peas and j balvin"
songinfo2020[47, 2] <- "Ariana Grande and Justin Bieber"

# 2021
songinfo2021[2, 1] <- "Montero Call Me by Your Name"
songinfo2021[10, 1] <- "Beggin"
songinfo2021[10, 2] <- "Maneskin"
songinfo2021[12, 1] <- "Dakiti"
songinfo2021[17, 2] <- "Silk Sonic"
songinfo2021[21, 2] <- "Tiesto"
songinfo2021[25, 2] <- "Riton and Nightcrawlers"
songinfo2021[25, 1] <- "Friday dopamine re edit"
songinfo2021[26, 1] <- "telepatia"
songinfo2021[33, 2] <- "Myke Towers and Juhn"
songinfo2021[34, 2] <- "Maneskin"
songinfo2021[46, 1] <- "Que Mas Pues"
songinfo2021[48, 1] <- "34 35"
songinfo2021[50, 1] <- "Pareja Del Ano"
songinfo2021[50, 2] <- "Sebastian yatra and Myke Towers"

# 2022
songinfo2022[5, 1] <- "Titi Me Pregunto"
songinfo2022[7, 1] <- "Enemy"
songinfo2022[8, 1] <- "quevedo bzrp music sessions vol 52"
songinfo2022[8, 2] <- "Bizarrap and quevedo"
songinfo2022[10, 1] <- "Running up that hill a deal with god"
songinfo2022[19, 2] <- "Elley Duhe"
songinfo2022[22, 2] <- "Rauw alejandro and chencho corleone"
songinfo2022[29, 2] <- "Lost frequencies and calum scott"
songinfo2022[36, 1] <- "I Aint Worried"
songinfo2022[39, 1] <- "Una Noche en Medellin"
songinfo2022[42, 2] <- "Bad Bunny & Rauw Alejandro"
songinfo2022[44, 2] <- "Tiesto"
```
Loop for scraping (for some reason loop randomly breaks if too many numbers)
```{r}
# 2018
final_df <- data.frame()
for(j in 48:50) {
  artist_name <- songinfo2018[j, 2]
  song_title <- songinfo2018[j, 1]
  df <- get_lyrics_search(artist_name = artist_name, song_title = song_title)
  df <- df$line
  df <- as.vector(df)
  df<- paste(df, collapse = " ") 
  final_df <- rbind(final_df, df)
}

# 2019
for(j in 48:50) {
  artist_name <- songinfo2019[j, 2]
  song_title <- songinfo2019[j, 1]
  df <- get_lyrics_search(artist_name = artist_name, song_title = song_title)
  df <- df$line
  df <- as.vector(df)
  df<- paste(df, collapse = " ") 
  final_df <- rbind(final_df, df)
}

# 2020
for(j in 47:50) {
  artist_name <- songinfo2020[j, 2]
  song_title <- songinfo2020[j, 1]
  df <- get_lyrics_search(artist_name = artist_name, song_title = song_title)
  df <- df$line
  df <- as.vector(df)
  df<- paste(df, collapse = " ") 
  final_df <- rbind(final_df, df)
}

# 2021
for(j in 48:50) {
  artist_name <- songinfo2021[j, 2]
  song_title <- songinfo2021[j, 1]
  df <- get_lyrics_search(artist_name = artist_name, song_title = song_title)
  df <- df$line
  df <- as.vector(df)
  df<- paste(df, collapse = " ")
  final_df <- rbind(final_df, df)
}

# 2022
final_df <- data.frame()
for(j in 44:50) {
  artist_name <- songinfo2022[j, 2]
  song_title <- songinfo2022[j, 1]
  df <- get_lyrics_search(artist_name = artist_name, song_title = song_title)
  df <- df$line
  df <- as.vector(df)
  df<- paste(df, collapse = " ")
  final_df <- rbind(final_df, df)
}

colnames(final_df)[1] <- "lyrics"
setwd("~/Desktop/Fall 2022/Data methods/Group Project")
save(final_df, file="lyrics_data.Rda")
```
Label column and add year
```{r}
load("~/Desktop/Fall 2022/Data methods/Group Project/lyrics_data.Rda")
final_df$num <- seq.int(from = 1, to = 250)
final_df$year <- 2018
final_df$year[which(final_df$num > 50 & final_df$num < 101)] <- 2019
final_df$year[which(final_df$num > 100 & final_df$num < 151)] <- 2020
final_df$year[which(final_df$num > 150 & final_df$num < 201)] <- 2021
final_df$year[which(final_df$num > 200)] <- 2022
```
Paste all lyrics for year into one line of dataframe
https://stackoverflow.com/questions/54805201/how-to-paste-all-string-values-in-a-column-together-as-one 
```{r}
# insert year here
df2 <- final_df %>%
  group_by(year) %>%
  summarise(col = paste(lyrics, collapse=" "))
df2 <- df2 %>%
  ungroup()
write.csv(df2,"lyrics_oneline.csv", row.names = FALSE)
```
Text analysis
```{r}
count_result <- data.frame()
afinn_result <- data.frame()
bing_and_nrc_result <- data.frame()
for(j in 2018:2021) {
text <- df2$col[which(df2$year == j)]
tibble <- tibble(line = 1, text = text)
tidy_lyrics <- unnest_tokens(tibble, word, text)
data(stop_words)
tidy_lyrics <- tidy_lyrics %>%
  anti_join(stop_words)
# Count of common words
count <- tidy_lyrics %>%
  count(word, sort = TRUE) %>%
  mutate(year = j)
count_result <- rbind(count_result, count)
# Sentiment analysis using afinn
afinn <- get_sentiments("afinn")
afinn_text <- tidy_lyrics %>% 
  inner_join(afinn) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN") %>%
  mutate(year = j)
afinn_result <- rbind(afinn_result, afinn_text)
# Sentiment analysis using bing and nrc
bing_and_nrc <- bind_rows(
  tidy_lyrics %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
   tidy_lyrics %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative) %>%
  mutate(year = j)
bing_and_nrc_result <- rbind(bing_and_nrc_result, bing_and_nrc)
}

```
Get audio features
```{r}
year_id <- data.frame(id = c("37i9dQZF1DXe2bobNYDtW8", "37i9dQZF1DWVRSukIED0e9", "37i9dQZF1DX7Jl5KP2eZaS", "5GhQiRkGuqzpWZSE7OU4Se", "37i9dQZF1DX18jTM2l2fJY"), year = 2018:2022)

songs_ana <- data.frame()
for(j in 1:5) {
playlist_id <- year_id[j, 1]
year <- year_id[j, 2]
tracks <- get_playlist_audio_features(
  "spotify",
  playlist_id) %>%
  mutate(year = year)
songs_ana <- rbind(songs_ana, tracks)
}
songs_ana <- inner_join(songs_ana, top_songs, by = c("track.id", "year"))
songs_ana <- songs_ana %>%
  select(danceability, energy, key, loudness, speechiness, acousticness, instrumentalness, liveness, valence, tempo, year, track.name.y, track.id)
save(songs_ana, file="playlist audio features.Rda")
# write.csv(songs_ana,"playlist audio features.csv", row.names = FALSE)
```
Get lyrics for songs specifically about COVID-19
```{r}
covid_songs <- read_excel("covid songs.xlsx") %>%
  as.matrix()
final_df2 <- data.frame()
for(j in 7:23) {
  artist_name <-covid_songs[j, 2]
  song_title <- covid_songs[j, 1]
  df <- get_lyrics_search(artist_name = artist_name, song_title = song_title)
  df <- df$line
  df <- as.vector(df)
  df<- paste(df, collapse = " ") 
  final_df2 <- rbind(final_df2, df)
}
colnames(final_df2) <- "lyrics"
final_df2 <- final_df2 %>%
  summarise(col = paste(lyrics, collapse=" "))
save(final_df2, file="covid songs.Rda")
```






